Model-based variable selection for regression with highly correlated variables.

### 2014-08-21 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@808 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

{\bf Abstract.} Linear regression outcomes (estimates, prevision) are known to be damaged by highly correlated covariates. However most modern datasets are expected to mechanically convey more and more highly correlated covariates due to the global increase of the amount of variables they contain. We propose to explicitly model such correlations by a family of linear regressions between the covariates. The structure of correlations is found with an {\sc mcmc} algorithm aiming at optimizing a specific {\sc bic} criterion. This hierarchical-like approach leads to a joint probability distribution on both the initial response variable and the linearly explained covariates. Then, marginalisation on the linearly explained covariates produces a parsimonious correlation-free regression model from which classical procedures for estimating regression coefficient, including any variable selection procedures, can be plugged.
Both simulated and real-life datasets from steel industry, where correlated variables are frequent, highlight that this proposed covariable pretreatment-like method has two essential beneficits: First, it offers a real readability of the linear links between covariates; Second, it improves significantly efficiency of classical estimation/selection methods which are performed after.
An {\sc r} package (\textsc{CorReg}), available on the {\sc cran}, implements this new method.
\smallskip

@@

\section{Introduction}\label{sec:intro}
%la régression et ses problèmes
Linear regression is a very standard and efficient method providing a predictive model with a good interpretability even for non-statisticians. Therefore, it is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}: Astronomy \cite{isobe1990linear}, Sociology \cite{longford2012revision}, Industry (real datasets of the present paper), \dots
With the rise of informatics, datasets contain more and more covariates leading high variance estimates, misleading interpretations and poor prediction accuracy for two different, quite related, reasons. First, the number of covariates leads to more complex models. Second, the number of covariates mechanically increases the chance to have correlated ones. Both situations are quite intricate and appear technically in a similar manner in the estimate procedure through ill-conditioned matrices. As a consequence, explicit break-up between them is not very common, most proposed methods focusing on the seminal question of the number of covariates selection. Originality of this paper is to distinguish explicitly them by proposing a specific decorrelation step followed by any classical variable selection step chosen by the practitioner. The decorrelation step can be viewed as a variable selection step but it is focused only on correlations, not on the number of covariables.

@@

Reducing the variance induced by the large number of covariables can be reached by targeting a better bias-variance trade off. Relating methods are numerous and continue to generate a lot of work. Most traditional directions are shrinkage, including variable selection, and also variable clustering.

Ridge regression \cite{marquardt1975ridge} is a shrinkage method which proposes possibly biased estimator that can be written in terms of a parametric $L_2$ penalty. It does not select variables since coefficients tend to 0 but don't reach 0, leading to difficult interpretations for a large number of covariates.	Since real datasets may imply many irrelevant variables, variable selection should be prefered for more interpretable models. Variable selection methods may add also some bias by deleting some relevant covariates but it may reduced drastically the variance by the dimension reduction.
As an emblematic method, the Least Absolute Shrinkage and Selection Operator (\textsc{lasso} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a parametric $L_1$ penalty to shrink some coefficients exactly to zero. But, like the ridge regression, the penalty does not distinguish correlated and independent covariates so there is no guarantee to have less correlated covariates. It only produces a parsimonious model, that is a gain for interpretation but only half the way from our point of view. In particular, \textsc{lasso} is also known to face consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates. So the quality of interpretation is compromised. Elastic net \cite{zou2005regularization} is a method developed to reach a compromise between Ridge regression and the \textsc{lasso} by a linear combination of $L_1$ and $L_2$ penalties. But, since it is based on the grouping effect, correlated covariates get similar coefficients and are selected together.

Another way for improving the conditioning and the understandability is to consider clusters of variables with the same coefficients, like the Octogonal Shrinkage and Clustering Algorithm for Regression (\textsc{oscar} \cite{bondell2008simultaneous}) to reduce dimension and also correlations if correlated covariates are in the same clusters. A possible bias is added by the dimension reduction inherent to the coefficients clustering.
The CLusterwise Effect REgression (\textsc{clere} \cite{yengo2012variable}) describes the regression coefficients no longer as fixed effect parameters but as unobserved independent random variables with grouped coefficients following a Gaussian Mixture distribution. The idea is that if the model has a small number of groups of covariates, the model will have a number of free parameters significantly lower than the number of covariates. In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on the coefficients estimator. 
Spike and Slab variable selection \cite{ishwaran2005spike} also relies on a Gaussian mixture (the spike and the slab) hypothesis for the regression coefficients and gives a subset of covariates (not grouped) on which to compute the Ordinary Least Squares estimate (\textsc{ols}) but has no specific protection against correlations issues.

@@

The idea of the present paper is that if we know explicitly the correlations, we could use this knowledge to avoid this specific problem. Correlations are thus new information to reduce the variance without adding any bias. More precisely, correlations are modeled through a system of linear sub-regressions between covariates. The set of covariates which are {\it never} at the place of a response variable in these sub-regressions is finally the greatest set of orthogonal covariates. Marginalizing over the dependent co-variables leads then to a linear regression (in relation to the initial response variable) with only orthogonal covariates. This marginalization step can be viewed also as a variable selection step but guided only by the correlations between covariates. Advantages of this approach is twofold. First, it improves interpretation through a good readability of dependency between covariables. Second, this marginal model is still a ``true'' model provided that both the initial regression model and all the sub-regressions are ``true''. As a consequence, the associated {\sc ols} will preserve an unbiased estimate but with a possibly reduced variance comparing to the {\sc ols} with the full regression model. The fact that the variance decreases depends on the residual variances involved in the sub-regressions: The more the sub-regressions are marked, the less will be the variance of associated {\sc ols}. In fact, any other estimation method than {\sc ols} can be plugged after the marginalization step. Indeed, it can be viewed as a pretreatment against correlation which can be chained after with dimension reduction methods, without no more suffering from correlations this time.
The sub-linear structure is obtained by a {\sc mcmc} algorithm optimizing a specific {\sc bic} criterion associated to the joint distribution on the covariates, regardless of the initial response variable. This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{cran}.

@@

\subsection{Sub-regressions between covariates}%\label{sectionOLS}
The classical linear regression model can be written
\begin{equation}
		\boldsymbol{Y}{|\boldsymbol{X}};\boldsymbol{\beta},\sigma_Y^2 = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
	\end{equation}
with $\boldsymbol{X}=(\boldsymbol{X}^1,\ldots,\boldsymbol{X}^d)$ the $n\times d$ matrix of $d$ predictor variables $\boldsymbol{X}^j$ ($j=1,\ldots,d$), $\boldsymbol{Y}$ the $n\times 1$ vector of the $n$ response variables and $\boldsymbol{\varepsilon} \sim \mathcal{N}_n(\boldsymbol{0},\sigma_Y^2\boldsymbol{I})$ the centered Gaussian noise of the regression with standard deviation $\sigma_Y >0$, $\boldsymbol{I}$ denoting the identity matrix of suitable dimension. The $d\times 1$ vector $\boldsymbol{\beta}$ gathers the coefficients of the regression\footnote{Usually a constant is included as one of the regressors. For example we can take $\boldsymbol{X}^1=(1,\ldots,1)'$. The corresponding element of $\boldsymbol{\beta}$ is then the intercept $\beta_1$.}, that can be estimated by Ordinary Least Squares (\textsc{ols}): %As shown in section \ref{sectionOLS}, 
	\begin{equation}
		\hat{\boldsymbol{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y} \label{eq:OLS}
	\end{equation}
	It is an unbiased estimate with variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\boldsymbol{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}. \label{eq:varOLS}
	\end{equation}
	This estimate requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which can lead to great variance estimates, so unstable estimates, if it is ill-conditioned. Ill-conditioning increases when the number $d$ of covariates grows and/or when correlations within the covariates grows (in absolute value)  \cite{hoerl1970ridge}. At the limit, when $d>n$ and/or when some correlations are maximum, $\boldsymbol{X}'\boldsymbol{X}$ becomes singular. Note that $d$ and correlations are also two non unrelated factors since, in real applications, increasing $d$ makes the risk to obtain correlated covariables higher.

@@

\begin{hyp}\label{H2}
In addition, we make the complementary hypothesis that the responses covariables and the predictor covariables are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariables and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates. This new assumption allows to obtain very simple sub-regressions sequences, discarding hierarchical ones, in particular uninteresting cyclic sub-regressions. However it is not too much restrictive since any hierarchical (but non-cyclic) sequence of sub-regressions can be agglomerated into a non-hierarchical sequence of sub-regressions, even if it may implies to partially loose information through variance increase in the new non-hierarchical sub-regressions.
\end{hyp}

\paragraph{Further notations} In the following, we will note also $\boldsymbol{J}_r=(J_{r}^1,\ldots,J_r^{d_r})$ the $d_r$-uple of all the response variable (to be not confused with the corresponding set $J_r$ previously defined), $\boldsymbol{J}_p=(J_{p}^1,\ldots,J_p^{d_r})$ the $d_r$-uple of all the predictors for all the sub-regressions, $\boldsymbol{d}_p=(d_p^1,\ldots,d_p^{d_{r}})$ the associated number of predictors and $\boldsymbol{S}=(\boldsymbol{J}_r,\boldsymbol{J}_p)$ the global {\it model} of all the sub-regressions. As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including these ones used as predictor covariables in $\boldsymbol{J}_p$ or not. An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}. The parameters are also stacked together: $\boldsymbol{\alpha}=(\boldsymbol{\alpha}_1,\ldots,\boldsymbol{\alpha}_{d_r})$ denotes the global coefficient of sub-regressions and $\boldsymbol{\sigma}^2=(\boldsymbol{\sigma}^2_1,\ldots,\boldsymbol{\sigma}^2_{d_r})$ denotes the corresponding global variance.

\paragraph{Remarks}
\begin{itemize}
\item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
\item As a consequence of Hypothesis~\ref{H1}, ``free'' covariables $\boldsymbol{X}_f$ are {\it all} decorrelated.
\item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals. Thus $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$ and it can be estimated independently as we will see in Section~\ref{sec:estimateS}.
\item The model of sub-regressions $\boldsymbol{S}$ gives a system of linear regressions that can be viewed as a recursive Simultaneous Equation Model (\textsc{sem})\cite{davidson1993estimation,TIMM} or also as a Seemingly Unrelated Regression (\textsc{sur}) \cite{SURzellner}. 
\end{itemize}

@@

Decomposing $\boldsymbol{\beta}=(\boldsymbol{\beta}_r',\boldsymbol{\beta}_f')'$ by $\boldsymbol{\beta}_r=\boldsymbol{\beta}_{J_r}$ and $\boldsymbol{\beta}_f=\boldsymbol{\beta}_{J_f}$ the regression coefficients associated respectively to the response and to the free covariates, we can rewrite (\ref{regressionsimple}):
\begin{equation}
			\boldsymbol{Y}{|\boldsymbol{X},\boldsymbol{S}};\boldsymbol{\beta},\sigma_Y^2=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon}_Y. \label{eq:MainR}
\end{equation}
Combining now (\ref{eq:MainR}) with (\ref{eq:SR}), (\ref{eq:marginal}) and (\ref{eq:H3}), and also independence between each $\boldsymbol{\varepsilon}_j$ and $\boldsymbol{\varepsilon}$, we obtain the following closed-form for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$:
\begin{eqnarray}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl}
\end{eqnarray}
Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariables $\boldsymbol{X}_f$. This decorrelation process has also acted like a specific covariable selection process-like because $\boldsymbol{X}_f \subseteq \boldsymbol{X}$. These two statements are expected to decrease the variance of further estimates of $\boldsymbol{\beta}$. However, the counterpart is twofold. First, this regression has a higher residual variance than the initial one since it is now $\sigma^{2*}_Y=\sigma^2_Y+\sum_{j \in I_r}\beta_{j}^2\sigma^2_j$ instead of $\sigma^2_Y$. Second, covariable selection being equivalent to set $\hat{\boldsymbol{\beta}}_r=\boldsymbol{0}$, it implies possibly biased estimates of $\boldsymbol{\beta}_r$. As a conclusion, we are faced to a typical {\it bias-variance trade off}. We will illustrate it in the next section in the case of the {\sc ols} estimate.

@@

\paragraph{Numerical illustration} We now illustrate the bias-variance trade off, through a numerical example. Let a simple case with $d=5$ covariables $\boldsymbol{X}=(\boldsymbol{X}^1,\boldsymbol{X}^2, \boldsymbol{X}^4, \boldsymbol{X}^5)$ independently drawn by four independent Gaussian $\mathcal{N}_n(\boldsymbol{0},\boldsymbol{I})$. Thus, $J_f=\{1,2,4,5\}$ and $\boldsymbol{X}_f=(\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{X}^4,\boldsymbol{X}^5)$. Let also $d_r=1$ response covariate $\boldsymbol{X}^3|\boldsymbol{X}^1,\boldsymbol{X}^2=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{\varepsilon}_1$ where $\boldsymbol{\varepsilon}_1\sim{\mathcal{N}_n(\boldsymbol{0},\sigma_1^2\boldsymbol{I})}$. Thus, $\boldsymbol{\alpha}_1=(1,1)'$, $\boldsymbol{J}_r=(3)$, $J_r=\{3\}$, $\boldsymbol{X}_r=(\boldsymbol{X}^3)$, $\boldsymbol{d}_p=(2)$, $\boldsymbol{J}_p=(\{1,2\})$, $\boldsymbol{X}^{\boldsymbol{J}_p^1}=(\boldsymbol{X}^1,\boldsymbol{X}^2)$ and $\boldsymbol{S}=((3),(\{1,2\}))$. Concerning now the regression with $\boldsymbol{Y}$, we define $\boldsymbol{Y}|\boldsymbol{X}=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{X}^3+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_Y$, where $\boldsymbol{\beta}=(1,1,1,1,1)'$ and $\sigma_Y \in \{10,20\}$. Finally, we deduce that $\boldsymbol{Y}|\boldsymbol{X}_f,\boldsymbol{S}=2\boldsymbol{X}^1+2\boldsymbol{X}^2+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_1+\boldsymbol{\varepsilon}_Y$
It is clear that $\boldsymbol{X}'\boldsymbol{X}$ will become more ill-conditioned as $\sigma_1$ gets smaller.

@@

\paragraph{Comparisons 2 to 4} We compare now three variable selections methods: {\sc lasso} with {\sc CorReg}+{\sc lasso}, elasticnet with {\sc CorReg}+electicnet and also stepwise with {\sc CorReg}+stepwise. Figures \ref{toutlar}, \ref{toutelast} and \ref{toutstepwise} respectively display the results on the same manner as the previous {\sc ols} with {\sc CorReg}+{\sc ols}. We see that {\sc CorReg}, used as a pre-treatment, provides similar prediction accurary as the three variable selection methods (this prediction is often better for small datasets) but with much more parsimonious regression models on $\boldsymbol{Y}$. This is remarkable because the true model depends on all covariates, so it highlight that other variable selection methods may be really penalized by the correlations between the covariates.

>>>>

{\bf Abstract.} Linear regression outcomes (estimates, prevision) are known to be damaged by highly correlated covariates. However most modern datasets are expected to mechanically convey more and more highly correlated covariates due to the global increase of the amount of variables they contain. We propose to explicitly model such correlations by a family of linear regressions between the covariates. The structure of correlations is found with an {\sc mcmc} algorithm aiming at optimizing a specific {\sc bic} criterion. This hierarchical-like approach leads to a joint probability distribution on both the initial response variable and the linearly explained covariates. Then, marginalisation on the linearly explained covariates produces a parsimonious correlation-free regression model from which classical procedures for estimating regression coefficient, including any variable selection procedures, can be plugged.
Both simulated and real-life datasets from steel industry, where correlated variables are frequent, highlight that this proposed covariates pretreatment-like method has two essential benefits: First, it offers a real readability of the linear links between covariates; Second, it improves significantly efficiency of classical estimation/selection methods which are performed after.
An {\sc r} package (\textsc{CorReg}), available on the {\sc cran}, implements this new method.
\smallskip

@@

\section{Introduction}\label{sec:intro}
%la régression et ses problèmes
Linear regression is a very standard and efficient method providing a predictive model with a good interpretability even for non-statisticians. Therefore, it is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}: Astronomy \cite{isobe1990linear}, Sociology \cite{longford2012revision}, Industry (real datasets of the present paper), \dots
With the rise of informatics, datasets contain more and more covariates leading high variance estimates, misleading interpretations and poor prediction accuracy for two different, quite related, reasons. First, the number of covariates leads to more complex models. Second, the number of covariates mechanically increases the chance to have correlated ones. Both situations are quite intricate and appear technically in a similar manner in the estimate procedure through ill-conditioned matrices. As a consequence, explicit break-up between them is not very common, most proposed methods focusing on the seminal question of the number of covariates selection. Originality of this paper is to distinguish explicitly them by proposing a specific decorrelation step followed by any classical variable selection step chosen by the practitioner. The decorrelation step can be viewed as a variable selection step but it is focused only on correlations, not on the number of covariates.

@@

Reducing the variance induced by the large number of covariates can be reached by targeting a better bias-variance trade off. Relating methods are numerous and continue to generate a lot of work. Most traditional directions are shrinkage, including variable selection, and also variable clustering.

Ridge regression \cite{marquardt1975ridge} is a shrinkage method which proposes possibly biased estimator that can be written in terms of a parametric $L_2$ penalty. It does not select variables since coefficients tend to 0 but don't reach 0, leading to difficult interpretations for a large number of covariates.	Since real datasets may imply many irrelevant variables, variable selection should be preferred for more interpretable models. Variable selection methods may add also some bias by deleting some relevant covariates but it may reduced drastically the variance by the dimension reduction.
As an emblematic method, the Least Absolute Shrinkage and Selection Operator (\textsc{lasso} \cite{tibshirani1996regression}) consists in a shrinkage of the regression coefficients based on a parametric $L_1$ penalty to shrink some coefficients exactly to zero. But, like the ridge regression, the penalty does not distinguish correlated and independent covariates so there is no guarantee to have less correlated covariates. It only produces a parsimonious model, that is a gain for interpretation but only half the way from our point of view. In particular, \textsc{lasso} is also known to face consistency problems \cite{Zhao2006MSC} when confronted with correlated covariates. So the quality of interpretation is compromised. Elastic net \cite{zou2005regularization} is a method developed to reach a compromise between Ridge regression and the \textsc{lasso} by a linear combination of $L_1$ and $L_2$ penalties. But, since it is based on the grouping effect, correlated covariates get similar coefficients and are selected together.

Another way for improving the conditioning and the understandability is to consider clusters of variables with the same coefficients, like the Octagonal Shrinkage and Clustering Algorithm for Regression (\textsc{oscar} \cite{bondell2008simultaneous}) to reduce dimension and also correlations if correlated covariates are in the same clusters. A possible bias is added by the dimension reduction inherent to the coefficients clustering.
The CLusterwise Effect REgression (\textsc{clere} \cite{yengo2012variable}) describes the regression coefficients no longer as fixed effect parameters but as unobserved independent random variables with grouped coefficients following a Gaussian Mixture distribution. The idea is that if the model has a small number of groups of covariates, the model will have a number of free parameters significantly lower than the number of covariates. In such a case, it improves interpretability and ability to yeld reliable prediction with a smaller variance on the coefficients estimator. 
Spike and Slab variable selection \cite{ishwaran2005spike} also relies on a Gaussian mixture (the spike and the slab) hypothesis for the regression coefficients and gives a subset of covariates (not grouped) on which to compute the Ordinary Least Squares estimate (\textsc{ols}) but has no specific protection against correlations issues.

@@

The idea of the present paper is that if we know explicitly the correlations, we could use this knowledge to avoid this specific problem. Correlations are thus new information to reduce the variance without adding any bias. More precisely, correlations are modeled through a system of linear sub-regressions between covariates. The set of covariates which are {\it never} at the place of a response variable in these sub-regressions is finally the greatest set of orthogonal covariates. Marginalizing over the dependent co-variables leads then to a linear regression (in relation to the initial response variable) with only orthogonal covariates. This marginalization step can be viewed also as a variable selection step but guided only by the correlations between covariates. Advantages of this approach is twofold. First, it improves interpretation through a good readability of dependency between covariates. Second, this marginal model is still a ``true'' model provided that both the initial regression model and all the sub-regressions are ``true''. As a consequence, the associated {\sc ols} will preserve an unbiased estimate but with a possibly reduced variance comparing to the {\sc ols} with the full regression model. The fact that the variance decreases depends on the residual variances involved in the sub-regressions: The more the sub-regressions are marked, the less will be the variance of associated {\sc ols}. In fact, any other estimation method than {\sc ols} can be plugged after the marginalization step. Indeed, it can be viewed as a pretreatment against correlation which can be chained after with dimension reduction methods, without no more suffering from correlations this time.
The sub-linear structure is obtained by a {\sc mcmc} algorithm optimizing a specific {\sc bic} criterion associated to the joint distribution on the covariates, regardless of the initial response variable. This algorithm is part of the R package \textsc{CorReg} accessible on \textsc{cran}.

@@

\subsection{Sub-regressions between covariates}%\label{sectionOLS}
The classical linear regression model can be written
\begin{equation}
		\boldsymbol{Y}{|\boldsymbol{X}};\boldsymbol{\beta},\sigma_Y^2 = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
	\end{equation}
with $\boldsymbol{X}=(\boldsymbol{X}^1,\ldots,\boldsymbol{X}^d)$ the $n\times d$ matrix of $d$ predictor variables $\boldsymbol{X}^j$ ($j=1,\ldots,d$), $\boldsymbol{Y}$ the $n\times 1$ vector of the $n$ response variables and $\boldsymbol{\varepsilon} \sim \mathcal{N}_n(\boldsymbol{0},\sigma_Y^2\boldsymbol{I})$ the centered Gaussian noise of the regression with standard deviation $\sigma_Y >0$, $\boldsymbol{I}$ denoting the identity matrix of suitable dimension. The $d\times 1$ vector $\boldsymbol{\beta}$ gathers the coefficients of the regression\footnote{Usually a constant is included as one of the regressors. For example we can take $\boldsymbol{X}^1=(1,\ldots,1)'$. The corresponding element of $\boldsymbol{\beta}$ is then the intercept $\beta_1$.}, that can be estimated by Ordinary Least Squares (\textsc{ols}): %As shown in section \ref{sectionOLS}, 
	\begin{equation}
		\hat{\boldsymbol{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y} \label{eq:OLS}
	\end{equation}
	It is an unbiased estimate with variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\boldsymbol{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}. \label{eq:varOLS}
	\end{equation}
	This estimate requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which can lead to great variance estimates, so unstable estimates, if it is ill-conditioned. Ill-conditioning increases when the number $d$ of covariates grows and/or when correlations within the covariates grows (in absolute value)  \cite{hoerl1970ridge}. At the limit, when $d>n$ and/or when some correlations are maximum, $\boldsymbol{X}'\boldsymbol{X}$ becomes singular. Note that $d$ and correlations are also two non unrelated factors since, in real applications, increasing $d$ makes the risk to obtain correlated covariates higher.

@@

\begin{hyp}\label{H2}
In addition, we make the complementary hypothesis that the responses covariates and the predictor covariates are totally disjoint: for any sub-regression $j=1,\ldots,d_{r}$, $J_{p}^j\subset J_f$ where $J_{r}=\{J_{r}^1,\ldots,J_{r}^{d_r}\}$ is set of all response covariates and $J_f=\{1,\ldots,d\} \backslash J_{r}$ is the set of all {\it non} response covariates. This new assumption allows to obtain very simple sub-regressions sequences, discarding hierarchical ones, in particular uninteresting cyclic sub-regressions. However it is not too much restrictive since any hierarchical (but non-cyclic) sequence of sub-regressions can be agglomerated into a non-hierarchical sequence of sub-regressions, even if it may implies to partially loose information through variance increase in the new non-hierarchical sub-regressions.
\end{hyp}

\paragraph{Further notations} In the following, we will note also $\boldsymbol{J}_r=(J_{r}^1,\ldots,J_r^{d_r})$ the $d_r$-uple of all the response variable (to be not confused with the corresponding set $J_r$ previously defined), $\boldsymbol{J}_p=(J_{p}^1,\ldots,J_p^{d_r})$ the $d_r$-uple of all the predictors for all the sub-regressions, $\boldsymbol{d}_p=(d_p^1,\ldots,d_p^{d_{r}})$ the associated number of predictors and $\boldsymbol{S}=(\boldsymbol{J}_r,\boldsymbol{J}_p)$ the global {\it model} of all the sub-regressions. As more compact notations, we define also $\boldsymbol{X}_r=\boldsymbol{X}^{J_{r}}$ the whole set of response covariates and also $\boldsymbol{X}_f=\boldsymbol{X}^{J_{f}}$ the {\it all} other covariates, denominating now as {\it free} covariates, including these ones used as predictor covariates in $\boldsymbol{J}_p$ or not. An illustration of all these notations is displayed through an example in Section~\ref{sec:tradeoff}. The parameters are also stacked together: $\boldsymbol{\alpha}=(\boldsymbol{\alpha}_1,\ldots,\boldsymbol{\alpha}_{d_r})$ denotes the global coefficient of sub-regressions and $\boldsymbol{\sigma}^2=(\boldsymbol{\sigma}^2_1,\ldots,\boldsymbol{\sigma}^2_{d_r})$ denotes the corresponding global variance.

\paragraph{Remarks}
\begin{itemize}
\item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
\item As a consequence of Hypothesis~\ref{H1}, ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
\item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals. Thus $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$ and it can be estimated independently as we will see in Section~\ref{sec:estimateS}.
\item The model of sub-regressions $\boldsymbol{S}$ gives a system of linear regressions that can be viewed as a recursive Simultaneous Equation Model (\textsc{sem})\cite{davidson1993estimation,TIMM} or also as a Seemingly Unrelated Regression (\textsc{sur}) \cite{SURzellner}. 
\end{itemize}

@@

Decomposing $\boldsymbol{\beta}=(\boldsymbol{\beta}_r',\boldsymbol{\beta}_f')'$ by $\boldsymbol{\beta}_r=\boldsymbol{\beta}_{J_r}$ and $\boldsymbol{\beta}_f=\boldsymbol{\beta}_{J_f}$ the regression coefficients associated respectively to the response and to the free covariates, we can rewrite (\ref{regressionsimple}):
\begin{equation}
			\boldsymbol{Y}{|\boldsymbol{X},\boldsymbol{S}};\boldsymbol{\beta},\sigma_Y^2=\boldsymbol{X}_f\boldsymbol{\beta}_f+\boldsymbol{X}_r\boldsymbol{\beta}_r+\boldsymbol{\varepsilon}_Y. \label{eq:MainR}
\end{equation}
Combining now (\ref{eq:MainR}) with (\ref{eq:SR}), (\ref{eq:marginal}) and (\ref{eq:H3}), and also independence between each $\boldsymbol{\varepsilon}_j$ and $\boldsymbol{\varepsilon}$, we obtain the following closed-form for the distribution of $\{\boldsymbol{Y} |\boldsymbol{X}_f,\boldsymbol{S};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}\}$:
\begin{eqnarray}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma^2}&=&\boldsymbol{X}_f (\boldsymbol{\beta}_f+ \sum_{j \in I_r}\beta_{j}\boldsymbol{\alpha}_j)+  \sum_{j \in I_r}\beta_{j}\boldsymbol{\varepsilon}_j+\boldsymbol{\varepsilon}_Y \label{eq:Trueexpl} \\
	&=&\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl}
\end{eqnarray}
Consequently, we have obtained a new regression expression of $\boldsymbol{Y}$ but relying now {\it only} on uncorrelated covariates $\boldsymbol{X}_f$. This decorrelation process has also acted like a specific variable selection process-like because $\boldsymbol{X}_f \subseteq \boldsymbol{X}$. These two statements are expected to decrease the variance of further estimates of $\boldsymbol{\beta}$. However, the counterpart is twofold. First, this regression has a higher residual variance than the initial one since it is now $\sigma^{2*}_Y=\sigma^2_Y+\sum_{j \in I_r}\beta_{j}^2\sigma^2_j$ instead of $\sigma^2_Y$. Second, variable selection being equivalent to set $\hat{\boldsymbol{\beta}}_r=\boldsymbol{0}$, it implies possibly biased estimates of $\boldsymbol{\beta}_r$. As a conclusion, we are faced with a typical {\it bias-variance trade off}. We will illustrate it in the next section in the case of the {\sc ols} estimate.

@@

\paragraph{Numerical illustration} We now illustrate the bias-variance trade off, through a numerical example. Let a simple case with $d=5$ covariates $\boldsymbol{X}=(\boldsymbol{X}^1,\boldsymbol{X}^2, \boldsymbol{X}^4, \boldsymbol{X}^5)$ independently drawn by four independent Gaussian $\mathcal{N}_n(\boldsymbol{0},\boldsymbol{I})$. Thus, $J_f=\{1,2,4,5\}$ and $\boldsymbol{X}_f=(\boldsymbol{X}^1,\boldsymbol{X}^2,\boldsymbol{X}^4,\boldsymbol{X}^5)$. Let also $d_r=1$ response covariate $\boldsymbol{X}^3|\boldsymbol{X}^1,\boldsymbol{X}^2=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{\varepsilon}_1$ where $\boldsymbol{\varepsilon}_1\sim{\mathcal{N}_n(\boldsymbol{0},\sigma_1^2\boldsymbol{I})}$. Thus, $\boldsymbol{\alpha}_1=(1,1)'$, $\boldsymbol{J}_r=(3)$, $J_r=\{3\}$, $\boldsymbol{X}_r=(\boldsymbol{X}^3)$, $\boldsymbol{d}_p=(2)$, $\boldsymbol{J}_p=(\{1,2\})$, $\boldsymbol{X}^{\boldsymbol{J}_p^1}=(\boldsymbol{X}^1,\boldsymbol{X}^2)$ and $\boldsymbol{S}=((3),(\{1,2\}))$. Concerning now the regression with $\boldsymbol{Y}$, we define $\boldsymbol{Y}|\boldsymbol{X}=\boldsymbol{X}^1+\boldsymbol{X}^2+\boldsymbol{X}^3+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_Y$, where $\boldsymbol{\beta}=(1,1,1,1,1)'$ and $\sigma_Y \in \{10,20\}$. Finally, we deduce that $\boldsymbol{Y}|\boldsymbol{X}_f,\boldsymbol{S}=2\boldsymbol{X}^1+2\boldsymbol{X}^2+\boldsymbol{X}^4+\boldsymbol{X}^5+\boldsymbol{\varepsilon}_1+\boldsymbol{\varepsilon}_Y$
It is clear that $\boldsymbol{X}'\boldsymbol{X}$ will become more ill-conditioned as $\sigma_1$ gets smaller.

@@

\paragraph{Comparisons 2 to 4} We compare now three variable selections methods: {\sc lasso} with {\sc CorReg}+{\sc lasso}, elasticnet with {\sc CorReg}+electicnet and also stepwise with {\sc CorReg}+stepwise. Figures \ref{toutlar}, \ref{toutelast} and \ref{toutstepwise} respectively display the results on the same manner as the previous {\sc ols} with {\sc CorReg}+{\sc ols}. We see that {\sc CorReg}, used as a pre-treatment, provides similar prediction accuracy as the three variable selection methods (this prediction is often better for small datasets) but with much more parsimonious regression models on $\boldsymbol{Y}$. This is remarkable because the true model depends on all covariates, so it highlight that other variable selection methods may be really penalized by the correlations between the covariates.
